Explore Vegetation Data

### 2016-02-15 explore_veg_data.Rmd
Update untagged species DBH data.
Made more generalized data set of alternate species DBH and individual stem and plot-levels. I assigned sampling years of 2005 and 2014, corresponding to plot establishment and the most recent remeasurement.

Now I need to examine the data for alternate and non-host species DBH. These are the stems that are untagged and meet a 5cm DBH threshold.

```{r untagged species}
summary(alt_sp_dbh)
library(lubridate)
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
filter(alt_sp_dbh, is.na(date)) # this should be 5/22/2012
alt_sp_dbh$Date[alt_sp_dbh$Date == ""] <- "5/22/2012 0:00:00"
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
summary(alt_sp_dbh)
alt_sp_dbh$year <- year(alt_sp_dbh$date)
alt_sp_dbh <- alt_sp_dbh %>% select(-PlotVisit_PlotVisitID, -Date) %>% 
      rename(dbh = Measure)
unique(as.character(alt_sp_dbh$Species))

alt_sp_dbh$Species[alt_sp_dbh$Species == "PSME"] <- "psme"
alt_sp_dbh$Species <- tolower(alt_sp_dbh$Species)
summary(alt_sp_dbh)
unique(alt_sp_dbh$Species)
unique(alt_sp_dbh$Dead)
# alt_sp_dbh$Dead <- ifelse(alt_sp_dbh$Dead == 0, "no", "yes")
alt_sp_dbh <- as.tbl(alt_sp_dbh)

summary(filter(alt_sp_dbh, year == 2003))
unique(droplevels(filter(alt_sp_dbh, year == 2003)$PlotID))
pid2003 <- as.character(unique(filter(alt_sp_dbh, year == 2003)$PlotID))
alt_sp_2003 <- filter(alt_sp_dbh, year == 2003)

summary(filter(alt_sp_dbh, year == 2004))
unique(droplevels(filter(alt_sp_dbh, year == 2004)$PlotID))
pid2004 <- as.character(unique(filter(alt_sp_dbh, year == 2004)$PlotID))
alt_sp_2004 <- filter(alt_sp_dbh, year == 2004)

@@

summary(filter(alt_sp_dbh, year == 2005))
unique(droplevels(filter(alt_sp_dbh, year == 2005)$PlotID))
pid2005 <- as.character(unique(filter(alt_sp_dbh, year == 2005)$PlotID))
alt_sp_2005 <- filter(alt_sp_dbh, year == 2005)

@@

filter(alt_sp_45, PlotID == "LINS01")
filter(alt_sp_2003, PlotID == "LINS01")
# Keep LINS01 from 2003 data, add to other data
filter(alt_sp_45, PlotID == "GANAY02")
filter(alt_sp_2003, PlotID == "GANAY02")
# Exclude GANAY02 from the 2004, keep the 2003 data
alt_sp_45 <- filter(alt_sp_45, PlotID != "GANAY02")
unique(alt_sp_2003$PlotID) %in% unique(alt_sp_45$PlotID)
alt_sp_45 <- rbind(alt_sp_45, alt_sp_2003)
alt_sp_2005 <- alt_sp_45
rm(alt_sp_45)

@@

# Create data frames of live and dead stems for '2005' and '2014' samplings. In reality the 2005 grouping has data from visits during 2003 and 2004, and the 2014 grouping includes data from some visits in 2012.
alt_sp_2014_live <- filter(alt_sp_2014, Dead == 0)
alt_sp_2014_live <- select(alt_sp_2014_live, -Dead)
alt_sp_2005_live <- filter(alt_sp_2005, Dead == 0)
alt_sp_2005_live <- select(alt_sp_2005_live, -Dead)
alt_sp_2014_dead <- filter(alt_sp_2014, Dead == 1)
alt_sp_2014_dead <- select(alt_sp_2014_dead, -Dead)
alt_sp_2005_dead <- filter(alt_sp_2005, Dead == 1)# only 12 obs, all at GANAY02
alt_sp_2005_dead <- select(alt_sp_2005_dead, -Dead)

@@

```{r make wide format species abundance}
unique(alt_sp_2014_live$Species)
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="arcto sp."] <- "arcto"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="prunus sp."] <- "prunus"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="unknown sp. 1"] <- "unk1"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="quercus sp."] <- "qusp"

>>>>

# Untagged Overstory
Now I need to examine the data for alternate and non-host species DBH. These are the stems that are untagged and meet a 5cm DBH threshold.

```{r untagged species}
summary(alt_sp_dbh)
library(lubridate)
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
filter(alt_sp_dbh, is.na(date)) # this should be 5/22/2012
alt_sp_dbh$Date[alt_sp_dbh$Date == ""] <- "5/22/2012 0:00:00"
alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
summary(alt_sp_dbh)
alt_sp_dbh$year <- year(alt_sp_dbh$date)
head(alt_sp_dbh)

alt_sp_dbh <- alt_sp_dbh %>% select(-PlotVisit_PlotVisitID, -Date) %>% 
      rename(dbh = Measure)
unique(as.character(alt_sp_dbh$Species))

alt_sp_dbh$Species[alt_sp_dbh$Species == "PSME"] <- "psme"
alt_sp_dbh$Species <- tolower(alt_sp_dbh$Species)
summary(alt_sp_dbh)
unique(alt_sp_dbh$Species)
unique(alt_sp_dbh$Dead)
unique(alt_sp_dbh$year)
# alt_sp_dbh$Dead <- ifelse(alt_sp_dbh$Dead == 0, "no", "yes")
alt_sp_dbh <- as.tbl(alt_sp_dbh)

summary(filter(alt_sp_dbh, year == 2003))
unique(droplevels(filter(alt_sp_dbh, year == 2003)$PlotID))# 32 plots
pid2003 <- as.character(unique(filter(alt_sp_dbh, year == 2003)$PlotID))
alt_sp_2003 <- filter(alt_sp_dbh, year == 2003)

summary(filter(alt_sp_dbh, year == 2004))
unique(droplevels(filter(alt_sp_dbh, year == 2004)$PlotID))# 120 plots
pid2004 <- as.character(unique(filter(alt_sp_dbh, year == 2004)$PlotID))
alt_sp_2004 <- filter(alt_sp_dbh, year == 2004)

@@

summary(filter(alt_sp_dbh, year == 2005))
unique(droplevels(filter(alt_sp_dbh, year == 2005)$PlotID))# 25 plots
pid2005 <- as.character(unique(filter(alt_sp_dbh, year == 2005)$PlotID))
alt_sp_2005 <- filter(alt_sp_dbh, year == 2005)

@@

filter(alt_sp_45, PlotID == "LINS01")
filter(alt_sp_2003, PlotID == "LINS01")
# Keep LINS01 from 2003 data, add to other data
filter(alt_sp_45, PlotID == "GANAY02")
filter(alt_sp_2003, PlotID == "GANAY02")
# Exclude GANAY02 from the 2004, keep the 2003 data
alt_sp_45 <- filter(alt_sp_45, PlotID != "GANAY02")
unique(alt_sp_2003$PlotID) %in% unique(alt_sp_45$PlotID)# LINS01 still there
alt_sp_45 <- rbind(alt_sp_45, alt_sp_2003)
alt_sp_2005 <- alt_sp_45
rm(alt_sp_45)

@@

alt_species <- rbind(alt_sp_2005, alt_sp_2014)
write.csv(alt_species, "analysis/data/alt-species-dbh-corrected.csv", row.names = F)
```

```{r group alt dbh to plot and year}
alt_species
alt_species$Dead[alt_species$Dead==0] <- "Alive"
alt_species$Dead[alt_species$Dead==1] <- "Dead"
alt_species <- rename(alt_species, status = Dead, plotid = PlotID, species = Species)
alt_dbh_plot <- alt_species %>% 
      group_by(plotid, species, status, year) %>% 
      summarise(tot_dbh = sum(dbh, na.rm = T), avg_dbh = mean(dbh, na.rm = T),
                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
                avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T),
                abundance = length(dbh))
write.csv(alt_dbh_plot, "analysis/data/alt-dbh-plot.csv", row.names = F)
```

```{r }
# Create data frames of live and dead stems for '2005' and '2014' samplings. In reality the 2005 grouping has data from visits during 2003 and 2004, and the 2014 grouping includes data from some visits in 2012.
alt_sp_2014_live <- filter(alt_sp_2014, Dead == 0)
alt_sp_2014_live <- select(alt_sp_2014_live, -Dead)
alt_sp_2005_live <- filter(alt_sp_2005, Dead == 0)
alt_sp_2005_live <- select(alt_sp_2005_live, -Dead)
alt_sp_2014_dead <- filter(alt_sp_2014, Dead == 1)
alt_sp_2014_dead <- select(alt_sp_2014_dead, -Dead)
alt_sp_2005_dead <- filter(alt_sp_2005, Dead == 1)# only 12 obs, all at GANAY02
alt_sp_2005_dead <- select(alt_sp_2005_dead, -Dead)

@@

```{r make wide format species abundance}
alt_species$PlotID <- tolower(alt_species$PlotID)
unique(alt_species$Species)
alt_species$Species[alt_species$Species=="arcto sp."] <- "arcto"
alt_species$Species[alt_species$Species=="prunus sp."] <- "prunus"
alt_species$Species[alt_species$Species=="unknown sp. 1"] <- "unk1"
alt_species$Species[alt_species$Species=="quercus sp."] <- "qusp"
alt_species

unique(alt_sp_2014_live$Species)
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="arcto sp."] <- "arcto"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="prunus sp."] <- "prunus"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="unknown sp. 1"] <- "unk1"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="quercus sp."] <- "qusp"
